---
title: "DataTransfomation"
author: "Yi Chen"
date: "9/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(nycflights13)
library(gapminder)
library(Lahman)
#tidyverse_update()
```

## Data Transformation
```{r}
flights <- tibble(flights)
class(flights)
str(flights)
```


### Filter

near() is a safe wat of compare if two vector of floading point numbers are equal.

#### Examples
```{r}
# Logical operators 
filter(flights, month == 1, day == 1) %>% head(5)

filter(flights, month %in% c(11,12) & (dep_time >400 | sched_dep_time < 600) & day != 1)

flights %>% filter(between(hour,0,6))%>% select(1:5,time_hour) %>% head(3)

# Missing values
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
filter(df, is.na(x) | x > 1)
```
#### Exercises
```{r}
# Had an arrival delay of two or more hours
flights %>% filter(arr_delay >=2) %>% head(3)

# Flew to Houston (IAH or HOU)
flights %>% filter(dest=="IAH" | dest=="HOU") %>% select(1:5,dest) %>% head(3)

# Were operated by United, American, or Delta
flights %>% filter(carrier %in% c("AA","UA","DL"))%>% select(1:5,carrier) %>% head(3)

# Arrived more than two hours late, but didn’t leave late
flights %>% filter(arr_delay>2 & dep_delay<=0)%>% select(1:5,dep_delay, arr_delay) %>% head(3)

# Were delayed by at least an hour, but made up over 30 minutes in flight
flights %>% filter(dep_delay>1 & air_time>30)%>% select(1:5,hour) %>% head(3)


# Departed between midnight and 6am (inclusive)
flights %>% filter(hour<=6)%>% select(1:5,time_hour) %>% head(3)

#Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
flights %>% filter(between(hour,0,6))%>% select(1:5,time_hour) %>% head(3)
filter(flights, between(month, 7, 9))


#How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
flights %>% filter(is.na(dep_time))

#Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
NA^0
NA*0
FALSE & NA
```


### Order/Arrange
#### Examples
```{r}
flights %>% arrange(year,month,day)

flights %>% arrange(desc(dep_delay))%>% select(1:5,dep_delay) %>% head(3)


# Missing values are always sorted at the end (no matter desc or asec)
df %>% arrange(x)
df %>% arrange(desc(x))

```
#### Exercises

```{r}
#How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
flights %>% arrange(desc(is.na(.)))
flights %>% arrange(desc(is.na(dep_time)))

#Sort flights to find the most delayed flights. Find the flights that left earliest.
flights %>% arrange(desc(dep_delay)) %>% slice(1)
flights %>% arrange(dep_delay)%>% slice(1)

#Sort flights to find the fastest (highest speed) flights.
flights %>% arrange(desc(distance/air_time)) %>% top_n(1)

flights %>% mutate(speed=distance/air_time) %>%arrange(desc(speed))  %>% top_n(1)

#Which flights travelled the farthest? Which travelled the shortest?
flights %>% arrange()
```

### slice
#### Examples
```{r}
# first the first 5 results
flights %>% top_n(5)
flights %>% head(5)

# find the first 10 results but obmit the first three
flights %>% slice(4:10)
flights %>% filter(between(row_number(), 4, 10))

# find the first 1e-05 records
flights %>% top_frac(1e-05)
```

### Select
Tidyverse selections implement a dialect of R where operators make it easy to select variables:

* : for selecting a range of consecutive variables.

* ! for taking the complement of a set of variables.

* & and | for selecting the intersection or the union of two sets of variables.

* c() for combining selections.

In addition, you can use *selection helpers*. Some helpers select specific columns:

* everything(): Matches all variables.

```{r}
## move some colums to the begining
select(flights, time_hour, air_time, everything())
```


* last_col(): Select last variable, possibly with an offset.

These *helpers* select variables by matching patterns in their names:

* starts_with(): Starts with a prefix.

* ends_with(): Ends with a suffix.

* contains(): Contains a literal string.

* matches(): Matches a regular expression.

* num_range(): Matches a numerical range like x01, x02, x03.

These *helpers* select variables from a character vector:

* all_of(): Matches variable names in a character vector. All names must be present, otherwise an out-of-bounds error is thrown.

* any_of(): Same as all_of(), except that no error is thrown for names that don't exist.

This *helper* selects variables with a function:

* where(): Applies a function to all variables and selects those for which the function returns TRUE.
```{r}
flights %>% select(where(is.character))

iris %>% select(where(~is.numeric(.x) && mean(.x) > 3.5))
```

#### Examples
```{r}
flights %>% select(year, month, day)
flights %>% select(year:day)
flights %>% select(-(year:day))
```

#### Exercises
```{r}
#select dep_time, dep_delay, arr_time, and arr_delay from flights.
flights %>% select(dep_time, dep_delay, arr_time, arr_delay)
select(flights, matches("^(dep|arr)_(time|delay)$"))
select(flights, starts_with("dep_"), starts_with("arr_"))
variables <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
select(flights, !!variables)

#What happens if you include the name of a variable multiple times in a select() call?
flights %>% select(dep_time, dep_time)

#What does the one_of() function do? Why might it be helpful in conjunction with this vector?
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
flights %>% select(one_of(vars))


#Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
select(flights, contains("TIME"))
select(flights, contains("TIME", ignore.case = FALSE))
select(flights, matches(".TIME."))
```

### mutate

* +, -, *, /, ^,%/% (integer division) and %% (remainder), log(), etc., for their usual mathematical meanings

* lead(), lag()
```{r}
lag(1:5)
lag(1:5, n = 2)
lag(1:5, default = 0)
lead(1:5)
lead(1:5,n=2)
lead(1:5, default = 6)
```


* dense_rank(), min_rank(), percent_rank(), row_number(), cume_dist(), ntile()
```{r}
x <- c(5, 1, 3, 2, 2, NA)
row_number(x)
min_rank(x)
dense_rank(x)
percent_rank(x)
cume_dist(x)

#a rough rank, which breaks the input vector into n buckets
ntile(x, 2)
ntile(1:8, 3)
```


* cumsum(), cummean(), cummin(), cummax(), cumany(), cumall()
```{r}
cumall(c(1,2,3,4,0))
cumany(c(1,2,3,4,0))
```

* na_if(), coalesce()
```{r}
# Use a single value to replace all missing values
x <- sample(c(1:5, NA, NA, NA))
coalesce(x, 0L)

# Or match together a complete vector from missing pieces
y <- c(1, 2, NA, NA, 5)
z <- c(NA, NA, 3, 4, 5)
coalesce(y, z)

# Supply lists by with dynamic dots
vecs <- list(
  c(1, 2, NA, NA, 5),
  c(NA, NA, 3, 4, 5)
)
coalesce(!!!vecs)

```


* if_else(), recode(), case_when()
```{r}
x <- c(-5:5, NA)
if_else(x < 0, NA_integer_, x)
if_else(x < 0, "negative", "positive", "missing")

# Unlike ifelse, if_else preserves types
x <- factor(sample(letters[1:5], 10, replace = TRUE))
ifelse(x %in% c("a", "b", "c"), x, factor(NA))
if_else(x %in% c("a", "b", "c"), x, factor(NA))
# Attributes are taken from the `true` vector,

x <- 1:50
case_when(
  x %% 35 == 0 ~ "fizz buzz",
  x %% 5 == 0 ~ "fizz",
  x %% 7 == 0 ~ "buzz",
  TRUE ~ as.character(x)
)


char_vec <- sample(c("a", "b", "c"), 10, replace = TRUE)
recode(char_vec, a = "Apple")
recode(char_vec, a = "Apple", b = "Banana")
```

#### Examples
```{r}
(flights_sml <- flights %>% select( 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time) %>% mutate(
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60,
  gain_per_hour = gain / air_time * 60))

## only keep the mutated variables
flights %>% transmute(gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours)
```

#### Exercises
```{r}
#Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
time2mins <- function(x) {
  (x %/% 100 * 60 + x %% 100) %% 1440
}
flights_times <- mutate(flights,
  dep_time_mins = time2mins(dep_time),
  sched_dep_time_mins = time2mins(sched_dep_time)
)
# view only relevant columns
select(
  flights_times, dep_time, dep_time_mins, sched_dep_time,
  sched_dep_time_mins
)

#Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
flights_delayed <- mutate(flights, 
                          dep_delay_min_rank = min_rank(desc(dep_delay)),
                          dep_delay_row_number = row_number(desc(dep_delay)),
                          dep_delay_dense_rank = dense_rank(desc(dep_delay))
                          )
flights_delayed <- filter(flights_delayed, 
                          !(dep_delay_min_rank > 10 | dep_delay_row_number > 10 |
                              dep_delay_dense_rank > 10))
flights_delayed <- arrange(flights_delayed, dep_delay_min_rank)

select(flights_delayed, month, day, carrier, flight, dep_delay, 
             dep_delay_min_rank, dep_delay_row_number, dep_delay_dense_rank)
```


### Aggregrate/Summary
#### Examples
```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))

flights %>% group_by( year, month, day) %>% summarise(delay = mean(dep_delay, na.rm = TRUE))
```

#### Exercises



### group by
#### Examples
#### Exercises


### Join/Merge
#### Examples
#### Exercises

### others

```{r}
rename(flights, tail_num = tailnum) %>% select(tail_num)
```
```{r}
# replace all na
df <- tibble(x = c(1, 2, NA), y = c("a", NA, "b"))
df %>% replace_na(list(x = 0, y = "unknown"))
```

