---
title: "exercise"
author: "Yi Chen"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# create tablesx
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)

(student <- tibble(sid=1:10,
                  s_name = c("zhao lei","qian dian","sun feng","li yun","zhou mei", "wu lan","zheng zhu","wang ju","zhao lei","zhao lei"),
                  s_birth = c(ymd(19900101),ymd(19901221),ymd(19900520),ymd(19900806),ymd(19911212),ymd(20171213),ymd(19890701),ymd(19900120),ymd(19900121),ymd(19900122)),
                  s_sex = c("Male","Male","Male","Male","Female","Female","Female","Female","Female","Male")))

(score <- tibble(s_id = c(rep(1:4,each=3),rep(5:7,each=2)),
                c_id = c(rep(1:3,4),1,3,1,3,3,1),
                s_score = c(80,90,99,70,60,rep(80,4),50,30,20,76,87,31,34,89,98)))

(course <- tibble(c_id=1:3,
                  c_name = c('chinese','math','english'),
                  t_id = c(2,1,3)))

(teacher <- tibble(t_id = 1:3,
                   t_name = c('zhang san','li si','wang wu')))
```

# questions

#### find the information about the students where students have higher score on course 1 than course 2

```{r}
(answer_1 <- score %>% 
   pivot_wider(names_from = c_id, values_from = s_score,names_prefix='class_') %>%
   select(s_id,class_1,class_2) %>%
   filter(!is.na(class_2) & class_1 > class_2) %>%
   left_join(student,by = c("s_id" = "sid"))%>%
   arrange(s_id))
```

**Mysql**

select st.s_id, st.s_name, sc1.s_score "course 1", sc2.s_score "course 2" 

from Student as st 

inner join Score as sc1 on sc1.s_id = st.s_id and sc1.c_id = "01" 

inner join Score sc2 on st.s_id = sc2.s_id and sc2.c_id = "02" 

where sc1.s_score < sc2.s_score;


#### find the information about the students where students have lower score on course 1 than course 2
```{r}
(answer_2 <- score %>%
   pivot_wider(names_from = c_id, values_from=s_score,names_prefix='class_') %>%
   select(s_id,class_1,class_2) %>%
   filter(!is.na(class_2) & class_1 < class_2) %>%
   left_join(student,by=c("s_id"="sid")) %>% 
   arrange(s_id))
```

#### find the students information (id, name, and average score) whoes average score is bigger than 60
```{r}
(answer_3 <- score %>%
   group_by(s_id) %>% 
   summarise(average_score = round(mean(s_score,na.rm=TRUE),2)) %>%
   filter(average_score > 60) %>%
   left_join(student,by = c("s_id"="sid")) %>%
   select(s_id,s_name,average_score) %>%
   arrange(s_id)
 )
```

#### find the information about students' id, name, number of course and total score
```{r}
(answer_4 <- score %>%
             group_by(s_id) %>% summarise(N_course = n(),total_score = sum(s_score)) %>%
             left_join(student,by = c("s_id"="sid")) %>%
             select(s_id,s_name,N_course,total_score) %>%
             arrange(s_id))
```

#### find the number of teacher whose family name is li
```{r}
(answer_5 <- teacher %>% 
             filter(grepl("^li",t_name)))

```


#### find the student information who has taken the class tought by zhang san
```{r}
(answer_6 <- score %>% left_join(course,by=c("c_id"="c_id")) %>%
             left_join(teacher,by=c("t_id"="t_id")) %>%
             filter(t_name == "zhang san") %>% 
             left_join(student,by= c("s_id"="sid")) %>%
             arrange(s_id))
```

#### find the student information who has not taken the class taught by zhang san
```{r}
(students_take_zhangsan <- score %>% left_join(course,by = c("c_id" = "c_id")) %>%
             left_join(teacher,by = c("t_id"="t_id")) %>%
             filter(t_name=="zhang san") %>%
             select(s_id))
(answer_7 <- student %>% filter(! sid %in% pull(students_take_zhangsan)))
## pull: take the tibble column into vector
```

#### find the student information who have taken both course 1 and course 2
```{r}
(answer_8 <- score %>% 
   pivot_wider(names_from = c_id, values_from = s_score, names_prefix="course_") %>%
   filter(!is.na(course_1) &!is.na(course_2)) %>%
   left_join(student,by = c("s_id" = "sid")) %>%
   select(s_id,s_name,s_birth,s_sex) %>%
   arrange(s_id))


(answer_8 <- score %>% mutate(both_1_2 = ifelse(c_id %in% c(1,2),1,0)) %>% 
               group_by(s_id) %>% 
               summarise(sum_both = sum(both_1_2)) %>%
               filter(sum_both==2) %>%
               select(s_id) %>%
               left_join(student,by=c("s_id"="sid")) %>%
               arrange(s_id))
```

#### find the student information who have taken course 1 but not course 2
```{r}
(answer_9 <- score %>% 
    pivot_wider(names_from = c_id,values_from = s_score,names_prefix="course_") %>%
    filter(!is.na(course_1) & is.na(course_2)) %>%
    select(s_id) %>%
    left_join(student,by = c("s_id" = "sid")) %>%
    arrange(s_id)
 )

contain_1 <- score %>%
      mutate(taken_1 = ifelse(c_id == 1,1,0)) %>%
      group_by(s_id) %>%
      summarise(filter_1 = sum(taken_1)) %>%
      filter(filter_1==1) %>%
      select(s_id) %>%
      pull()
not_contain_2 <- score %>%
      mutate(taken_2 = ifelse(c_id == 2,1,0)) %>%
      group_by(s_id) %>%
      summarise(filter_2 = sum(taken_2)) %>%
      filter(filter_2<1) %>%
      select(s_id) %>%
      pull()
answer <- intersect(contain_1,not_contain_2)
(answer_9 <- score %>%
      filter(s_id %in% answer) %>%
      select(s_id) %>%
      left_join(student,by = c("s_id" = "sid")) %>%
      arrange(s_id)
)
   
```

#### find the student information who have not taken all availble course
```{r}
taken_all_course <- score %>%
    pivot_wider(names_from = c_id, values_from = s_score,names_prefix="course_") %>%
    drop_na() %>%
    select(s_id) %>%
    pull()
(answer_10 <- score %>% 
      filter(! s_id %in% taken_all_course) %>%
      distinct(s_id) %>%
      left_join(student, by = c("s_id" = "sid")) %>%
      arrange(s_id)
)


(answer_10 <- score %>% 
      group_by(s_id) %>%
      summarise(n_course = n()) %>%
      filter(n_course<max(n_course)) %>%
      select(s_id ) %>%
      left_join(student, by = c("s_id" = "sid")) %>%
      arrange(s_id)
)
   
```

#### find the students information who have taken the at least one same course with the student has id equal to 1.
```{r}
(score %>% filter(c_id %in%
                 pull(score %>% filter(s_id == 1) %>% distinct(c_id))) %>%
   distinct(s_id) %>%
   left_join(student,by = c("s_id"="sid")) %>%
    arrange(s_id)
)
```


#### find the students information who have taken the exact same courses with student has id equal to 1.

*hard*

```{r}
## group_concat

(answer_10 <- score %>%
    group_by(s_id) %>%
    summarise(courses = toString(sort(unique(c_id)))) %>% #*****#
    filter(courses == courses[1]) %>%
    distinct(s_id) %>%
    left_join(student,by = c("s_id"="sid")) %>%
    arrange(s_id)
 ) 


```



#### find the student information who have ever taught by teacher zhang san.
```{r}
(answer_11 <- score %>% mutate(taught_by_zhangsan = 
              ifelse(c_id %in% 
              pull(teacher %>% filter(t_name =="zhang san") %>%
                  inner_join(course,by=c("t_id"="t_id"))%>%
                  select(c_id)),1,0)) %>%
    group_by(s_id) %>%
    summarise(N_zhangsan = sum(taught_by_zhangsan)) %>%
    filter(N_zhangsan==0) %>%
    distinct(s_id) %>%
   left_join(student,by = c("s_id"="sid")) %>%
    arrange(s_id))
```


#### find the student inforamtion (name and average score) who have failed in at least two courses.

```{r}
(answer_12 <- score %>% 
    mutate(fail = ifelse(s_score < 60,1,0)) %>%
    group_by(s_id) %>%
    summarise(N_fail = sum(fail),
              average_score = round(mean(s_score,na.rm = TRUE),2)) %>%
    filter(N_fail >=2) %>%
    left_join(student,by = c("s_id"="sid")) %>%
    arrange(s_id))
```
#### find the students who have taken course 1 and failed (score <60) and rank the student descreasingly by score
```{r}
(answer_13 <- score %>% 
    filter(c_id == 1 & s_score < 60) %>%
    left_join(student,by = c("s_id"="sid")) %>%
    arrange(desc(s_score))
)
```


#### list all information about the students (including the coures they token and corresponding score) and rank the student decreasingly by average score 
```{r}
(answer_14 <- score %>%
    left_join(course, by = c("c_id"="c_id")) %>%
    group_by(s_id) %>%
    summarise(course_taken = toString(sort(unique(c_name))),
              scores = toString(sort(unique(s_score))),
           average_score = mean(s_score,na.rm = TRUE)) %>%
    left_join(student, by = c("s_id" = "sid")) %>%
    arrange(desc(average_score))
 )

(avg_score <- score %>%
      group_by(s_id) %>%
      summarise(average_score = round(mean(s_score,na.rm =  TRUE),2)))
      
(answer_15 <- score %>% pivot_wider(names_from = c_id,values_from = s_score,names_prefix = "course_") %>%
      left_join(student, by = c("s_id" = "sid")) %>%
      left_join(avg_score, by = c("s_id" = "s_id")) %>%
      arrange(desc(average_score))
)


```

#### find the max, min, and average score for each course as well as the pass_rate
```{r}
(answer_16 <- score %>%
    group_by(c_id) %>%
    summarise(highest_score = max(s_score),
              lowest_score = min(s_score),
              pass_rate = round(mean(s_score>=60),2)) %>%
    left_join(course,by= c("c_id" = "c_id")) %>%
    arrange(desc(pass_rate))
    )
```

#### rank the score by each course and show the ranking

*hard*

```{r}
(answer_17 <- score %>%
    group_by(c_id) %>%
    mutate(ranking = rank(desc(s_score))) %>%
    arrange(c_id,ranking)
)


(answer_17 <- score %>%
    group_by(c_id) %>%
    mutate(ranking = min_rank(desc(s_score))) %>%
    arrange(c_id,ranking)
)

(answer_17 <- score %>%
    group_by(c_id) %>%
    mutate(ranking = dense_rank(desc(s_score))) %>%
    arrange(c_id,ranking)
)

(answer_17 <- score %>%
    group_by(c_id) %>%
    mutate(ranking = round(percent_rank(desc(s_score)),2)) %>%
    arrange(c_id,ranking)
)
```

#### find the total score for each student and rank them
```{r}
(answer_18 <- score %>%
    group_by(s_id) %>%
    summarise(total_score = sum(s_score)) %>%
    arrange(desc(total_score)) %>%
    left_join(student,by = c("s_id"="sid"))
)
```

#### find the average score for differnt teacher and different course and rank descreasingly by average score
```{r}
(answer_19 <- score %>%
    left_join(course,by=c("c_id"="c_id")) %>%
    left_join(teacher,by=c("t_id"="t_id")) %>%
    group_by(c_name,t_name) %>%
    summarise(average_score = round(mean(s_score,na.rm = TRUE),2)) %>%
    arrange(desc(average_score))
)
```
####  for each courese, find the 2th and 3th student's information
```{r}
(answer_20 <- score %>% 
    group_by(c_id) %>%
    mutate(ranking = dense_rank(s_score)) %>%
    arrange(c_id, s_score) %>%
    filter(ranking == 2 | ranking == 3) %>%
    left_join(student,by = c("s_id"="sid")) %>%
    left_join(course,by=c("c_id" = "c_id"))
)
```

#### find the proportation of student having score between (0-60),(60-90), and (90,100) for each course
```{r}
(answer_21 <- score %>%
    group_by(c_id) %>%
    summarise(fail_rate = round(sum(s_score<60)/n(),2),
              good_rate = round(sum(s_score>=60 & s_score<=90 )/n(),2),
              excellent_rate = round(sum(s_score>90)/n(),2))
)
```

#### find each students' average score and its ranking
```{r}
(answer_22 <- score %>%
   group_by(s_id) %>%
   summarise(average_score = round(mean(s_score),2)) %>%
   mutate(rank= dense_rank(desc(average_score))) %>%
   arrange(rank) %>%
   left_join(student,by=c("s_id"="sid"))
)
```

#### find the top three students for each course
```{r}
(answer_23 <- score %>%
    group_by(c_id) %>%
    mutate(rank = dense_rank(desc(s_score))) %>%
    arrange(c_id,rank) %>%
    filter(rank <= 3) 
    
)
```

#### find the number of students in each course
```{r}
(answer_24 <- score %>%
    group_by(c_id) %>%
    summarise(N_Student = n())
)
```

#### find the students who only take two courses
```{r}
(answer_25 <- score %>%
         group_by(s_id) %>%
         summarise(N_course = n()) %>%
         filter(N_course == 2) %>%
         left_join(student,by=c("s_id"="sid"))
)
```
#### find the number of male and female
```{r}
(answer_26 <- student %>%
    group_by(s_sex) %>%
    summarise(N_student=n())
)
```

#### find the student contains feng
```{r}
(answer_27 <- student %>%
    filter(grepl(".*feng.*",s_name))
 )
```

#### find the students who have the same name
```{r}
(answer_28 <- student %>%
   group_by(s_name) %>%
   summarise(N_student= n()) %>%
    filter(N_student>1)
)
```

#### find the student who born in 1990
```{r}
(answer_29 <- student %>%
    filter(year(s_birth)==1990)
)
```
#### find the average score for each course and rank them
```{r}
(answer_30 <- score %>%
    group_by(c_id) %>%
    summarise(average_score = round(mean(s_score),2)) %>%
    mutate(ranking = dense_rank(desc(average_score))) %>%
    arrange(ranking)
)
```
#### find all the students whoes average score is bigger than 85
```{r}
(answer_31 <- score %>%
    group_by(s_id) %>%
    summarise(average_score = round(mean(s_score),2)) %>%
    filter(average_score >=85) %>%
    left_join(student,by=c("s_id"="sid"))
 
 )
```
#### find the student in the math course whoes score is lower than 60
```{r}
(answer_32 <- score %>% 
    left_join(student,by = c("s_id"="sid")) %>%
    left_join(course,by= c("c_id" = "c_id")) %>%
    left_join(teacher, by = c("t_id"="t_id")) %>%
    filter(c_name == "math" & s_score > 80)
 
 )
```

#### for any courses, find the students whose socre is bigger than 70
```{r}
(answer_33 <- score %>%
    filter(s_score>70) %>%
    left_join(student,by = c("s_id"="sid")) %>%
    left_join(course,by= c("c_id" = "c_id"))
)
```

#### find all information of students who fail in any course
```{r}
(answer_34 <- score %>%
    filter(s_score<60) %>%
    left_join(student,by = c("s_id"="sid")) %>%
    left_join(course,by= c("c_id" = "c_id"))
 )
```

#### find the top student in zhang san's course

*hard*
```{r}
(answer_35 <- score %>% 
    left_join(student,by = c("s_id"="sid")) %>%
    left_join(course,by= c("c_id" = "c_id")) %>%
    left_join(teacher, by = c("t_id"="t_id")) %>%
    filter(t_name=="zhang san") %>%
    group_by(c_name) %>%
    filter(s_score==max(s_score))
)
```

#### find the same score from difffernt courses

*hard*
```{r}
(answer_36 <- score %>% 
    group_by(s_score) %>%
    filter(n()>1) %>%
    left_join(student,by = c("s_id"="sid")) %>%
    left_join(course,by= c("c_id" = "c_id"))
)
```



#### find the top two student from each course
```{r}
(anwer_37 <- score %>%
    group_by(c_id) %>%
    filter(dense_rank(desc(s_score))<=2) %>%
    arrange(c_id,desc(s_score))%>%
    left_join(student,by = c("s_id"="sid")) %>%
    left_join(course,by= c("c_id" = "c_id"))
)
```
#### count the numebr of student in each course (only care about the couse has at least five student)
```{r}
(answer_38 <- score %>%
    group_by(c_id) %>%
    summarise(n_student = n()) %>%
    filter(n_student > 5) %>%
    left_join(course,by= c("c_id" = "c_id"))
)
```

#### find the student who take at least two course
```{r}
(answer_39 <- score %>% 
    group_by(s_id) %>%
    filter(n()>=2) %>%
    left_join(student,by = c("s_id"="sid")) %>%
    left_join(course,by= c("c_id" = "c_id"))
)
```

#### find the age of student
```{r}
(anwer_38 <- student %>%
    mutate(age=year(now())-year(s_birth)))
```

#### find the students who have the birthday this week/month, next month
```{r}
(answer_40 <- student %>%
    filter(week(now()) == week(s_birth)))

(answer_40 <- student %>%
    filter(month(now()) == month(s_birth)))

(answer_40 <- student %>%
    filter(month(now())+1 == month(s_birth)))

```


#### find all combination of the two course
```{r}
courses <- course %>% distinct(c_name) %>% pull()
combination <- matrix(0,ncol=2,nrow = 1)
for (i in 1:(length(courses)-1)){
  for (j in (i+1):length(course)){
    print(c(i,j))
    combination <- rbind(combination,c(c(courses[i],courses[j])))
  }
}
combination[2:nrow(combination),]
```

#### find the serves time for each id
```{r}
data <- tibble(
  serves_id = c(1,1,2,2,3,3,4,4,4,4,5),
  status = c("r","o","r","o","r","o","r","o","r","o","r"),
  servise_date = c(ymd(20200101),ymd(20200108),ymd(20200105),ymd(20200111),ymd(20200110),ymd(20200122),ymd(20200202),ymd(20200204),ymd(20200210),ymd(20200216),ymd(20200301))
)
data <- data %>% group_by(serves_id) %>%
  mutate(row = row_number())
(answer <- data %>% 
    pivot_wider(names_from = status, values_from = servise_date,values_fill=ymd(today())) %>%
    mutate(diff_time = as.numeric(o-r)) %>%
    group_by(serves_id) %>%
    summarise(servise_days = sum(diff_time))
) 
  
```

