---
title: "string"
author: "Yi Chen"
date: "9/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## String basic (str_c: combine)
```{r message=FALSE, warning=FALSE}
library(tidyverse)

# strength length
str_length(c("a", "R for data science", NA))

# strength combine
str_c("x", "y", "z",sep = ',')

# str_c is vectorized and replace na with 'na'
x <- c("abc", NA)
str_c("|-", str_replace_na(x), "-|")
str_c("prefix-", c("a", "b", "c"), "-suffix")


# use if inside str_c
name <- "Hadley"
time_of_day <- "morning"
birthday <- TRUE

str_c(
  "Good ", time_of_day, " ", name,
  if (birthday) " and HAPPY BIRTHDAY",
  "."
)

# collapse
str_c(c("x", "y", "z"))
str_c(c("x", "y", "z"), collapse = ", ")

```

## Subsetting strings
```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
str_sub(x, -3, -1)


str_to_lower(str_sub(x, 1, 1))
str_to_upper(c("i", "ı"),locale = 'en')


str_sort(x, locale = "en")  # English
```

## regular expressions

* '.' matches any character;

* regexps use the backslash, \, to escape special behaviour;

* ^ to match the start of the string;

* $ to match the end of the string.

* \\d: matches any digit.

* \\s: matches any whitespace (e.g. space, tab, newline).

* \ [abc\ ]: matches a, b, or c.

* (a|b): matches a or b

* \ [^abc\ ]: matches anything except a, b, or c.

* ?: 0 or 1

* +: 1 or more

* *: 0 or more

* {n}: exactly n

* {n,}: n or more

* {,m}: at most m

* {n,m}: between n and m

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
str_view(x, ".a.")

```
1. (.)\\1\\1: The same character appearing three times in a row. E.g. "aaa"

2. "(.)(.)\\\\2\\\\1": A pair of characters followed by the same pair of characters in reversed order. E.g. "abba".

3. (..)\\1: Any two characters repeated. E.g. "a1a1".

4. "(.).\\\\1.\\\\1": A character followed by any character, the original character, any other character, the original character again. E.g. "abaca", "b8b.b".

5. "(.)(.)(.).*\\3\\2\\1" Three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order. E.g. "abcsgasgddsadgsdgcba" or "abccba" or "abc1cba".


### Detect matches
```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")

# How many common words start with t?
sum(str_detect(words, "^t"))

# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))

# select the elements that match a pattern
words[str_detect(words, "x$")]
str_subset(words, "x$")


# x <- c("apple", "banana", "pear")
str_count(x, "a")


tibble(
  word = words
)%>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

### Extract matches
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
(colour_match <- str_c(colours, collapse = "|"))
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)

(more <- sentences[str_count(sentences, colour_match) > 1])
str_view_all(more, colour_match)


tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  ) %>% head(5)
```


### Replacing matches
```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")

str_replace_all(x, "[aeiou]", "-")
```

### Splitting
```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]

#simplify = TRUE to return a matrix
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

