---
title: "apply family"
author: "Yi Chen"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## apply for array/matrix

apply(array,margin,function) ==> array in (n-1) dimension
```{r}
(x <- matrix(rnorm(30),ncol = 6))
apply(x,2,mean)
```


## lapply for dataframe, list, or vector

lapply(list,function) ==> list

```{r}
(x <- list(a = 1:10, beta = exp(-3:3), logic = c(TRUE,FALSE,FALSE,TRUE)))
lapply(x, mean)

A <- matrix(c(1,2,3,4,5,6,7,8,9),ncol = 3)
B <- matrix(c(4,5,6,7,8,9,10,11,12),ncol = 3)
C <- matrix(c(8,9,10,11,12,13),ncol = 2)
(mylist <- list(A,B,C))
lapply(mylist,"[",,2)
lapply(mylist,"[",1,)
lapply(mylist,"[",2,1)
```

## sapply function

similar to the lapply but to return the simplified list
```{r}
lapply(mylist,"[",2,1)
sapply(mylist, "[",2,1)
sapply(mylist,"[",2,1,simplify=FALSE)
unlist(lapply(mylist,"[",2,1))
```

## rep
```{r}
rep(c(1,2,3),c(3,2,1))
rep(1:4,len=21)
```

## mapply

multivariate apply: mapply applies FUN to the first elements of each ... argument, the second elements, the third elements, and so on.

```{r}
mapply(rep, 1:4, 4:1)


(Q1 <- matrix(c(rep(1, 4), rep(2, 4), rep(3, 4), rep(4, 4)),4,4))
mapply(rep,1:4,4)
```

## sweep
```{r}
require(stats) # for median
(med.att <- apply(attitude, 2, median))
head(sweep(data.matrix(attitude), 2, med.att),10)  # subtract the column medians


(A <- array(1:24, dim = 4:2))
sweep(A, 1:2, apply(A, 1:2, median))

```

## aggregrates


aggregate(x, ...)

* Default S3 method:

aggregate(x, ...)

* S3 method for class 'data.frame'

aggregate(x, by, FUN, ..., simplify = TRUE, drop = TRUE)

* S3 method for class 'formula'

aggregate(formula, data, FUN, ...,
          subset, na.action = na.omit)

* S3 method for class 'ts'

aggregate(x, nfrequency = 1, FUN = sum, ndeltat = 1,
          ts.eps = getOption("ts.eps"), ...)
          
```{r}
aggregate(state.x77, list(Region = state.region), mean)

aggregate(state.x77,
          list(Region = state.region,
               Cold = state.x77[,"Frost"] > 130),
          mean)


aggregate(weight ~ feed, data = chickwts, mean)
aggregate(breaks ~ wool + tension, data = warpbreaks, mean)
aggregate(cbind(Ozone, Temp) ~ Month, data = airquality, mean)
```

